Advancements in Moltres for Time-Dependent Multiphysics Molten Salt Reactor Modeling

### 2024-12-09 introduction.tex
Draft for committee before defense

    The second objective is to
    develop a novel hybrid $S_N$-diffusion method for accurate control rod modeling in
    time-dependent \gls{MSR} analyses. The hybrid method uses the discrete ordinates ($S_N$)
    neutron transport method to generate transport corrections in regions near control rods for
    drift correction terms in modified neutron diffusion equations. The $S_N$ method is applied on
    a reduced problem domain to ensure the hybrid method remains tractable on small to moderate
    computing clusters.
%\end{enumerate}

@@

Chapter 2 presents a literature review of existing multiphysics \gls{MSR} simulation software,
\gls{VV} studies on \gls{MSR} modeling and simulation, turbulence modeling in \gls{MSR} systems,
and transport-correction techniques for neutron diffusion methods.
Chapter 3 provides an in-depth description of Moltres and its existing capabilities in the context
of previously published work. This chapter then presents Moltres \gls{VV} results for the CNRS
benchmark and the numerical \gls{MSRE} zero-power pump experiment studies. Following these studies,
the chapter presents the implementation and verification of the Spalart-Allmaras turbulence model
in Moltres.
Chapter 4 presents the theory and numerical implementation of the hybrid
$S_N$-diffusion method in Moltres. These include the $S_N$ method implementation, the transport
correction formulations, the iteration algorithm, the $S_N$-diffusion coupling implementation, and
general implementation details relating to the underlying numerical solver in Moltres.
Chapter 5 presents the verification and demonstration of the hybrid $S_N$-diffusion method for
$k_\text{eff}$ and rod worth calculations through
$k$-eigenvalue simulations of 1-D, 2-D, and 3-D \gls{MSRE} models. The hybrid method is verified
against reference results from OpenMC Monte Carlo neutron transport code and \gls{MSRE}
experimental data.
Chapter 6 presents the demonstration of the hybrid $S_N$-diffusion method in time-dependent
reactivity-initiated simulations modeling the \gls{MSRE} rod drop and reactivity insertion
experiments.
Chapter 7 concludes this dissertation by summarizing the results presented, identifying limitations
in this work, and providing potential research directions to address those limitations or extend
this work.

>>>>

    The second objective is to develop, implement, and demonstrate a novel hybrid $S_N$-diffusion
    method for accurate control rod modeling in
    time-dependent \gls{MSR} analyses. The hybrid method uses the discrete ordinates ($S_N$)
    neutron transport method to generate transport corrections in regions near control rods for
    drift correction terms in modified neutron diffusion equations. The $S_N$ method is applied on
    a reduced problem domain to ensure the hybrid method remains tractable on small to moderate
    computing clusters.
%\end{enumerate}

@@

Chapter 2 presents a literature review of existing multiphysics \gls{MSR} simulation software,
\gls{VV} studies on \gls{MSR} modeling and simulation, turbulence modeling in \gls{MSR} systems,
and transport-correction techniques for neutron diffusion methods.
Chapter 3 provides an in-depth description of Moltres and its existing capabilities in the context
of previously published work. This chapter then presents Moltres \gls{VV} results for the CNRS
benchmark and the numerical \gls{MSRE} zero-power pump experiment studies. Following these studies,
the chapter presents the implementation and verification of the Spalart-Allmaras turbulence model
in Moltres.
Chapter 4 presents the theory and numerical implementation of the hybrid
$S_N$-diffusion method in Moltres. These include the $S_N$ method implementation, the transport
correction formulations, the iteration algorithm, the $S_N$-diffusion coupling implementation, and
general implementation details relating to the underlying numerical solver in Moltres.
Chapter 5 presents the verification and demonstration of the hybrid $S_N$-diffusion method for
$k_\text{eff}$ and rod worth calculations through
$k$-eigenvalue simulations of 1-D, 2-D, and 3-D \gls{MSRE} models. The hybrid method is verified
against reference results from OpenMC Monte Carlo neutron transport code and \gls{MSRE}
experimental data.
%Chapter 6 presents the demonstration of the hybrid $S_N$-diffusion method in time-dependent
%reactivity-initiated simulations modeling the \gls{MSRE} rod drop and reactivity insertion
%experiments.
Chapter 6 presents a demonstration of the hybrid $S_N$-diffusion method in a time-dependent
reactivity-initiated simulation modeling a \gls{MSRE} rod drop experiment.
Chapter 7 concludes this dissertation by summarizing the results presented, identifying limitations
in this work, and providing potential research directions to address those limitations or extend
this work.

### 2024-12-09 lit.tex
msre transient section additions
Draft for committee before defense

The existence of multiple, disparate physics in a single system constitutes a multiphysics system.
Consequently, we need multiphysics software to accurately model a multiphysics system. Employing
well-verified single-physics software for coupled multiphysics simulations does not guarantee
stability, accuracy, or robustness \cite{keyes_multiphysics_2013}. Nevertheless, coupling separate
single-physics software for multiphysics simulations is a viable and popular option with proper
implementation. With strong coupling between different physics, extra care must be
taken to capture various multiphysics interactions accurately. For instance, the strong coupling
between neutron flux and temperature in \glspl{MSR} in \glspl{MSR} necessitates robust coupling
techniques. Na\"ive multiphysics coupling algorithms may result in longer computation times
at best and a lack of convergence in simulations at worst.

@@

\subsection{Transport-Correction Techniques With Neutron Diffusion-Based Solvers}

@@

Methods based on absorber blackness \cite{davison_influence_1951, spinks_extrapolation_1965,
pellaud_extrapolation_1968, mendelson_two-dimensional_1969} encompass a
broad class of procedures for generating boundary conditions to match approximate solutions of
low-order methods (e.g., diffusion) to more accurate solutions from high-order methods (e.g.,
transport). The internal boundary conditions replace absorber regions and mimic their presence in
the problem domain. The boundary conditions are generalizations of the Marshak boundary condition
\cite{marshak_note_1947}, which in 1-D are of the form:
%
\begin{gather}
  \frac{\phi(x)}{d\phi(x)/dx} = \lambda \label{eq:marshak}
  \shortintertext{where}
  \begin{align}
    \lambda =& \mbox{ linear extrapolation length.} & \nonumber
  \end{align}
\end{gather}
%
The flux gradient in the denominator is taken in the outward direction at the surface of the
absorber or black body. In Cartesian coordinates, it is common to approximate Eq. \ref{eq:marshak}
in the form:
%
\begin{align}
  \phi(x+\lambda) =& 0
\end{align}
because Dirichlet boundary conditions are easier to work with analytically and numerically.
The corresponding relations in cylindrical and spherical coordinates depend on the radius of the
absorber rod. Various forms for $\lambda$ exist for specific absorber geometries such as slabs
\cite{maynard_blackness_1959} and cylinders \cite{spinks_extrapolation_1965,
pellaud_extrapolation_1968} which were dependent
on the size of the absorber region and coefficients quantifying the escape probability of neutrons
entering the absorber region. The $\lambda$ and the associated coefficients were derived
analytically from the \gls{BTE} with some simplifying assumptions, such as having uniform or
cosine-shaped incident neutron currents, isotropic scattering within the absorber, and mathematical
approximations in ignoring higher-order terms in intermediate steps. Blackness theory emerged in
the mid-20th century when computational resources were limited. Later, as computational
resources became more readily available and powerful enough for more complicated transport
calculations, appropriate boundary conditions could be calculated numerically
\cite{bretscher_computing_1997}. Alternatively, effective group constants can be calculated for
diffusion codes not programmed to handle internal boundary conditions. For instance,
Bretscher \cite{bretscher_computing_1997} provided formulae for effective diffusion coefficient and
absorption cross section for thin absorber slabs as a function of blackness coefficients, absorber
thickness, and mesh size.

@@

Lastly, there is a lack of research focus into robust techniques for accurate control rod modeling
for time-dependent \gls{MSR} analyses. Control rods induce highly anisotropic fluxes within
themselves and their vicinities.
Unfortunately, higher-fidelity neutron transport methods are typically too computationally
expensive for multiphysics \gls{MSR} simulations, while faster methods like neutron
diffusion perform poorly in regions with highly anisotropic fluxes. However, transport-correction
techniques can be applied to augment neutron diffusion calculations in these regions. While few
have been applied to \gls{MSR} multiphysics simulations other than albedo boundary conditions,
various techniques have been proposed and implemented for other reactor types. Helpful information
may be gleaned from these techniques to improve control rod modeling in \glspl{MSR} with the
neutron diffusion method.

>>>>

The existence of multiple, disparate physics in a single system constitutes a multiphysics system.
Consequently, we need multiphysics software to accurately model a multiphysics system. Employing
well-verified single-physics software for coupled multiphysics simulations does not guarantee
stability, accuracy, or robustness \cite{keyes_multiphysics_2013}. Nevertheless, coupling separate
single-physics software for multiphysics simulations is a viable and popular option with proper
implementation. With strong coupling between different physics, extra care must be
taken to capture various multiphysics interactions accurately. For instance, the strong coupling
between neutron flux and temperature in \glspl{MSR} necessitates robust coupling
techniques. Na\"ive multiphysics coupling algorithms may result in longer computation times
at best and a lack of convergence in simulations at worst.

@@

\subsection{Transport-Correction Techniques With Neutron Diffusion-Based Solvers}
\label{sec:lit-tc}

@@

Methods based on absorber blackness \cite{davison_influence_1951, spinks_extrapolation_1965,
pellaud_extrapolation_1968, mendelson_two-dimensional_1969} encompass a
broad class of procedures for generating boundary conditions to match approximate solutions of
low-order methods (e.g., diffusion) to more accurate solutions from high-order methods (e.g.,
transport). The internal boundary conditions replace absorber regions and mimic their presence in
the problem domain. The boundary conditions are generalizations of the Marshak boundary condition
\cite{marshak_note_1947}, which in 1-D are of the form:
%
\begin{gather}
  \frac{\phi(x)}{d\phi(x)/dx} = \lambda \label{eq:marshak}
  \shortintertext{where}
  \begin{align}
    \lambda =& \mbox{ linear extrapolation length.} & \nonumber
  \end{align}
\end{gather}
%
The flux gradient in the denominator is taken in the outward direction at the surface of the
absorber or black body. In Cartesian coordinates, it is common to approximate Eq. \ref{eq:marshak}
in the form:
%
\begin{align}
  \phi(x+\lambda) =& 0
\end{align}
because Dirichlet boundary conditions are easier to work with analytically and numerically.
The corresponding relations in cylindrical and spherical coordinates depend on the radius of the
absorber rod. Various forms for $\lambda$ exist for specific absorber geometries such as slabs
\cite{maynard_blackness_1959} and cylinders \cite{spinks_extrapolation_1965,
pellaud_extrapolation_1968} which were dependent
on the size of the absorber region and coefficients quantifying the escape probability of neutrons
entering the absorber region. $\lambda$ is derived
analytically from the \gls{BTE} with some simplifying assumptions, such as having uniform or
cosine-shaped incident neutron currents, isotropic scattering within the absorber, and mathematical
approximations in ignoring higher-order terms in intermediate steps. Absorber blackness methods
emerged in the mid-20th century when computational resources were limited. Later, as computational
resources became more readily available and powerful enough for more complicated transport
calculations, appropriate boundary conditions could be calculated numerically
\cite{bretscher_computing_1997}. Alternatively, effective group constants can be calculated for
diffusion codes not programmed to handle internal boundary conditions. For instance,
Bretscher \cite{bretscher_computing_1997} provided formulae for effective diffusion coefficient and
absorption cross section for thin absorber slabs as a function of blackness coefficients, absorber
thickness, and mesh size.

@@

Lastly, there is a lack of research focus into robust techniques for accurate control rod modeling
for time-dependent \gls{MSR} analyses. Control rods induce highly anisotropic fluxes within
themselves and their vicinities that the neutron diffusion approximation cannot model accurately.
Inaccurate rod worth estimates subsequently affect transient reactor power and temperature
distribution calculations. Higher-fidelity neutron transport methods remain too computationally
expensive for time-dependent multiphysics \gls{MSR} simulations. However, transport-correction
techniques can be applied to augment neutron diffusion calculations in these regions. While few
have been applied to \gls{MSR} multiphysics simulations other than albedo boundary conditions,
various techniques have been proposed and implemented for other reactor types. Helpful information
may be gleaned from these techniques to improve control rod modeling in \glspl{MSR} with the
neutron diffusion method.

### 2024-12-09 main.tex
add ray effect subsection
tikz preamble for previous commit
Draft for committee before defense

\usepackage[hyphens]{url}
\usepackage[hidelinks]{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hhline}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\allowdisplaybreaks
\usepackage{color}
\usepackage{multirow}
\usepackage{siunitx}
\usepackage{xfrac}
\usepackage{bm}
\usepackage{stmaryrd}

@@

\usepackage{tikz}
\usepackage{tkz-euclide}
\usetikzlibrary{positioning, arrows, decorations, shapes}
\usetikzlibrary{shapes.geometric,arrows}
\definecolor{illiniblue}{HTML}{B1C6E2}
\definecolor{illiniorange}{HTML}{f8c2a2}
\definecolor{green}{HTML}{c2e2b1}
\tikzstyle{process} = [rectangle, rounded corners, thick, minimum width=7cm, minimum height=1cm, text centered, draw=black, fill=illiniblue, text width=18em]
\tikzstyle{object} = [ellipse, minimum width=2cm, thick, minimum height=2.2cm, text centered, draw=black, fill=illiniorange, text width=12em]
\tikzstyle{decision} = [diamond, thick, aspect=2, minimum width=2cm, minimum height=2cm, text centered, draw=black, fill=green, text width=6em]
\tikzstyle{arrow} = [thick,->,>=stealth]

@@

%\title{A Hybrid Neutronics Method for Time-Dependent Control Rod Modeling in Molten Salt Reactors}
\title{Advancements in Moltres for Time-Dependent Multiphysics Molten Salt Reactor Modeling}
\author{Sun Myung Park}
\department{Nuclear, Plasma \& Radiological Engineering}
\degreeyear{2024}
\committee{Associate Professor Kathryn D. Huff, Chair \& Co-Director of Research \\
           Assistant Professor Madicken Munk, Co-Director of Research, Oregon State University \\
           Professor Rizwan Uddin \\
           Professor Tomasz Kozlowski \\
           Professor Paul Fischer}

>>>>

\usepackage[hyphens]{url}
\usepackage[hidelinks]{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[rightcaption]{sidecap}
\sidecaptionvpos{figure}{c}
\usepackage{hhline}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\allowdisplaybreaks
\usepackage{color}
\usepackage{multirow}
\usepackage{siunitx}
\usepackage{xfrac}
\usepackage{bm}
\usepackage{stmaryrd}
\usepackage{textcomp}

@@

\usepackage{tikz}
\usepackage{tkz-euclide}
\usetikzlibrary{positioning, arrows, decorations, shapes, fit, backgrounds}
\usetikzlibrary{shapes.geometric,arrows}
\definecolor{illiniblue}{HTML}{B1C6E2}
\definecolor{illiniorange}{HTML}{f8c2a2}
\definecolor{green}{HTML}{c2e2b1}
\tikzstyle{process} = [rectangle, rounded corners, thick, minimum width=7cm, minimum height=1cm, text centered, draw=black, fill=illiniblue, text width=18em]
\tikzstyle{object} = [ellipse, minimum width=2cm, thick, minimum height=2.2cm, text centered, draw=black, fill=illiniorange, text width=12em]
\tikzstyle{decision} = [diamond, thick, aspect=2, minimum width=2cm, minimum height=2cm, text centered, draw=black, fill=green, text width=6em]
\tikzstyle{arrow} = [thick,->,>=stealth]
\tikzstyle{solver} = [rectangle, rounded corners, thick, minimum width=3cm, minimum height=1.5cm, text centered, draw=black, fill=illiniblue, text width=7em]
\tikzstyle{bound} = [rectangle, rounded corners, thick, draw=black]

@@

%\title{A Hybrid Neutronics Method for Time-Dependent Control Rod Modeling in Molten Salt Reactors}
\title{Advancements in Moltres for Time-Dependent Multiphysics Molten Salt Reactor Modeling}
\author{Sun Myung Park}
\department{Nuclear, Plasma \& Radiological Engineering}
\degreeyear{2024}
\committee{Associate Professor Kathryn D. Huff, Co-Chair \\
           Assistant Professor Madicken Munk, Co-Chair \\
           Professor Rizwan Uddin \\
           Professor Tomasz Kozlowski \\
           Professor Paul Fischer}
